---
title: "age_metadata_analysis"
output: html_document
---

# create new age_years column in dataset21and22
```{r}
dataset21and22$age_years <- NA
dataset21and22$age_years <- as.numeric(dataset21and22$age_years)
```

# add "host age" data to "age_years" column
# create table for data not added to "age_years" -> "host_age_exceptions_df"
```{r}
column_name <- "host age"

# iterate by row
for (n_row in 1:nrow(dataset21and22)) {
  # check that host age has value
  if (!is.na(dataset21and22[n_row, column_name])) 
    {
    ### YEARS ###
    # check for "years" unit (y), excluding values with "-" and "day"
    if (grepl("y", dataset21and22[n_row, column_name], ignore.case=TRUE) &
        !grepl("-", dataset21and22[n_row, column_name]) &
        !grepl("day", dataset21and22[n_row, column_name], ignore.case=TRUE) &
        !grepl("m", dataset21and22[n_row, column_name], ignore.case=TRUE)) {
      dataset21and22[n_row, "age_years"] <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, column_name]))
    }
    # add exception "10 year-old"
    if (grepl("year-old", dataset21and22[n_row, column_name], ignore.case=TRUE)) {
      dataset21and22[n_row, "age_years"] <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, column_name]))
    }
    # add exception "0-1 years" (-> 1 y.o.)
    if (grepl("0-1 years", dataset21and22[n_row, column_name], ignore.case=TRUE)) {
      dataset21and22[n_row, "age_years"] <- 1
    }
    ### MONTHS ###
    if (grepl("m", dataset21and22[n_row, column_name], ignore.case=TRUE) &
        !grepl("y", dataset21and22[n_row, column_name], ignore.case=TRUE)) {
      # calculate months -> years
        age_months <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, column_name]))
        dataset21and22[n_row, "age_years"] <- round(age_months / 12)
    }
    ### DAYS ###
    if (grepl("day", dataset21and22[n_row, column_name], ignore.case=TRUE)) {
      # calculate days -> years
        age_days <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, column_name]))
        dataset21and22[n_row, "age_years"] <- round(age_days / 365)
    }
    if (grepl("D", dataset21and22[n_row, column_name])) {
      # calculate days -> years
        age_days <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, column_name]))
        dataset21and22[n_row, "age_years"] <- round(age_days / 365)
    }
    ### NEONATE EXCEPTION ###
    # add to "age_years" as 0 years
    if (grepl("neonate", dataset21and22[n_row, column_name], ignore.case=TRUE)) {
      dataset21and22[n_row, "age_years"] <- 0
    }
  }
} 

sum(!is.na(dataset21and22$age_years))
# host_age_exceptions_df <- subset(dataset21and22[,c(1,29)], !is.na(dataset21and22$`host age`) & is.na(dataset21and22$`age_years`))
```


# add "age" column data to "age_years" column
```{r}
# only exception is data with "days", so convert that to years and add to "age_years"
# other than that, assume all numbers are in years (bc why else...)
# iterate by row
for (n_row in 1:nrow(dataset21and22)) {
  # check that age_years hasn't been determined from host age
  if (is.na(dataset21and22[n_row, "age_years"])) {
    # check that "age" has value
    if (!is.na(dataset21and22[n_row, "age"])) {
      # check for values with "day" and convert to years
      if (grepl("day", dataset21and22[n_row, "age"])) {
        age_days <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, "age"]))
          dataset21and22[n_row, "age_years"] <- round(age_days / 365)
      }
      # directly transfer all other data from "age" to "age_years"
      else {
        dataset21and22[n_row, "age_years"] <- as.numeric(dataset21and22[n_row, "age"])
      }
        
    }
  }
}

```

### for "host age" w/o unit, check BioProject manually to confirm unit. Download list of BioSample ID's for each project, then add age to age_years with appropriate unit
```{r}
# sum of age_years values BEFORE = 667
# N samples w/ host age but not age_years = 1715 ; after list 0309 = 1563
# N samples w/ age_years = 3826 ; after list 0309 = 3978

# N samples w/ host age but not age_years = 1563 ; after list 0310 = 1445
# N samples w/ age_years = 3978 ; after list 0310 = 4096

# N samples w/ host age but not age_years = 1445 ; after list 0310_02 = 1389  
# N samples w/ age_years = 4096 ; after list 0310_02 = 4152

Bioproject = X0310_02_biosampleList_years
age_unit = "years"

# make list of samples in bioproject
sampleIDs <- as.vector(Bioproject[[1]])

n_y = 0
n_m = 0
n_d = 0

for (n_row in 1:nrow(dataset21and22)) {
  # check that there is a value in "host age"
  if (!is.na(dataset21and22[n_row,"host age"])) {
    #check that there is NOT a value in "age_years"
    if (is.na(dataset21and22[n_row,"age_years"])) {
      # check that value is "host age" is numeric (no dashes or words)
      if (!is.na(as.numeric(dataset21and22[n_row,"host age"]))) {
        # check if sample is in list of sample ID's for this bioproject
        if (dataset21and22[n_row, "sample_id"] %in% sampleIDs) {
          # if age unit is years, add "host age" directly to "age_years" column
          if (age_unit == "years") {
            dataset21and22[n_row, "age_years"] <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, "host age"]))
            n_y = n_y + 1
            # print(as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, "host age"])))
          }
          # if age unit is months, convert to years then add to "age_years"
          else if (age_unit == "months") {
            age_months <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, "host age"]))
            dataset21and22[n_row, "age_years"] <- round(age_months / 12)
            n_m = n_m + 1
          }
          # if age unit is days, convert to years then add to "age_years
          else if (age_unit == "days") {
            age_days <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, "host age"]))
            dataset21and22[n_row, "age_years"] <- round(age_days / 365)
            n_d = n_d + 1
          }
        
        # ONLY DO FOR AGE ASSUMPTION dataset21and22
        # check if age is reasonably in years (from 1-100)
        # if (as.numeric(dataset21and22[n_row,"host age"] >= 1 | as.numeric(dataset21and22[n_row,"host age"] <= 100))) {
        #   dataset21and22[n_row, "age_years"] <- as.numeric(gsub("[^0-9.]", "", dataset21and22[n_row, "host age"]))
        # }
      }
    }
  }
  }
}

# AFTER X01_biosampleList_days, sum of age_years values = 1484
# ages added = 817

# AFTER X01_biosampleList_years, sum of age_years values = 3487
# ages added = 2003

# AFTER X01=2_biosampleList_years, sum of age_years values = 3826
# ages added = 339

# AFTER age assumption, sum of age_years values = 4152
# ages added = 326

sum(!is.na(dataset21and22$age_years))
sum(subset(is.na(dataset21and22$age_years), !is.na(dataset21and22$"host age")))
host_age_exceptions_df <- subset(dataset21and22[,c(1,29)], !is.na(dataset21and22$`host age`) & is.na(dataset21and22$`age_years`))

# View(host_age_exceptions_df)
# View(table(host_age_exceptions_df$`host age`))

print(n_y)
print(n_m)
print(n_d)
```

# retrieve biosample ID's for samples w/ unconfirmed "host age" unit AND "host age" value is numeric (no dashes/ not a binned age range)
```{r}
# create empty vector to append biosample ID's to
biosample_ID_toCheckUnits <- list()

# iterate by row
for (n_row in 1:nrow(dataset21and22)) {
  # check that age_years is NA
  if (is.na(dataset21and22[n_row,"age_years"])) {
    # check that host age unit is numeric
    if (!is.na(as.numeric(dataset21and22[n_row,"host age"]))) {
      biosample_ID_toCheckUnits <- append(biosample_ID_toCheckUnits, dataset21and22[n_row,1])
    }
  }
}

# export as xlsx file
# write.xlsx(biosample_ID_toCheckUnits, file = "/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/dsrAB_Biosample_Metadata_Analysis/biosample_ID_list_toCheckUnits_0310_02.xlsx")

```

# add "host age" and "age_years" values to binned age ranges
# bin_9yrSize : binned ages in increments of 9 years (eg: 0-9, 10-19...)
# bin_5yrCutoff : binned ages with cutoff of 5 (<5, >=5)
```{r}
# create new columns for binned ages
subject_df$bin_9yrSize <- NA
subject_df$bin_5yrCutoff <- NA

# someone else's code to bin continuous numeric variable:
# 9 year bin size:
subject_df <- subject_df %>%
mutate(bin_9yrSize = cut(age_years, breaks = seq(0,110, by=10), include.lowest = TRUE, right = FALSE))

# 5 year cutoff: 
subject_df <- subject_df %>%
  mutate(bin_5yrCutoff = cut(age_years, breaks = c(0,5,110), right = FALSE))
# it works beautifully. thank god for the internet. mental note: stop solving problems in R by iterating by row bc it's super slow

# change to bin columns to character type AFTER binning age_years data
subject_df$bin_9yrSize <- as.character(subject_df$bin_9yrSize)
subject_df$bin_5yrCutoff <- as.character(subject_df$bin_5yrCutoff)

# must iterate by row to add binned ages for "host age" value exceptions
for (n_row in 1:nrow(subject_df)) {
  # skip over values with a value for age_years
  if (is.na(subject_df[n_row, "age_years"])) {
    # check if "host age" value !is.na
    if (!is.na(subject_df[n_row, "host age"])) {
      # determine bin values
      if (subject_df[n_row, "host age"] == "0-4" | 
          subject_df[n_row, "host age"] == "<5") {
        subject_df[n_row, "bin_9yrSize"] <- "[0,10)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[0,5)"
      }
      if (subject_df[n_row, "host age"] == "1 to 5") {
        subject_df[n_row, "bin_9yrSize"] <- "[0,10)"
      }
      if (subject_df[n_row, "host age"] == "5-9") {
        subject_df[n_row, "bin_9yrSize"] <- "[0,10)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "10-19") {
        subject_df[n_row, "bin_9yrSize"] <- "[10,20)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "20-29" |
          subject_df[n_row, "host age"] == "21 to 25") {
        subject_df[n_row, "bin_9yrSize"] <- "[20,30)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "30-39") {
        subject_df[n_row, "bin_9yrSize"] <- "[30,40)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "40-49") {
        subject_df[n_row, "bin_9yrSize"] <- "[40,50)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "50-59" |
          subject_df[n_row, "host age"] == "51 to 55") {
        subject_df[n_row, "bin_9yrSize"] <- "[50,60)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "60-69" |
          subject_df[n_row, "host age"] == "61 to 65") {
        subject_df[n_row, "bin_9yrSize"] <- "[60,70)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "70-79" |
          subject_df[n_row, "host age"] == "71 to 75") {
        subject_df[n_row, "bin_9yrSize"] <- "[70,80)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "80-89") {
        subject_df[n_row, "bin_9yrSize"] <- "[80,90)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      if (subject_df[n_row, "host age"] == "90-99") {
        subject_df[n_row, "bin_9yrSize"] <- "[90,100)"
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
      # don't fit into a 9 year bin
      if (subject_df[n_row, "host age"] == "26-30" |
          subject_df[n_row, "host age"] == "26-55 years" |
          subject_df[n_row, "host age"] == "35-44" |
          subject_df[n_row, "host age"] == "45-54" |
          subject_df[n_row, "host age"] == "55-64" |
          subject_df[n_row, "host age"] == "56 to 60" |
          subject_df[n_row, "host age"] == "65-74" |
          subject_df[n_row, "host age"] == "76 to 80" |
          subject_df[n_row, "host age"] == ">90") {
        subject_df[n_row, "bin_5yrCutoff"] <- "[5,110)"
      }
    }
  }
}

# SUCCESS!!! 

```

