---
title: "Biosample per Subject"
output: html_document
---

# Let's make subject_df  from dataset21and22!

```{r}
# first, let's take a look at the available metadata variables and single out the one's related to subject ID 

# View(colnames(dataset21and22))

variables_related_to_subject <- c("label",
                                  "sample name",
                                  "unique ID",
                                  "host subject id",
                                  "description",
                                  "Replication #",
                                  "Test person",
                                  "biospecimen repository",
                                  "submitted subject id",
                                  "patient_name",
                                  "Patient_name",
                                  "sample_alias",
                                  "ID",
                                  "patient",
                                  "individual name",
                                  "individual",
                                  "subject_id")

for (variable in variables_related_to_subject[17]){
  View(table(dataset21and22[,variable]))
}

# check out the samples
subset(dataset21and22$sample_id, dataset21and22$`subject_id` == "ERP005989|mother_64")

# unique ID possibly may be of interest -- need to look in to the studies
# should be used to identify single subjects:
subject_variables <- c(
  "host subject id",
  "Test person",
  "submitted subject id",
  "patient_name",
  "Patient_name",
  "patient",
  "individual name",
  "individual",
  "subject_id"
)

# now that we have a list of variables with subject information, we can add those samples to a df with row = subject ID
# first, let's figure out how many samples have overlapping data for the variables, and decide which variable should have priority when parsing the data

```

```{r}
# first, let's figure out how many samples have overlapping data for the variables, and decide which variable should have priority when parsing the data

# sum(!is.na(dataset21and22$`host subject id`) & !is.na(dataset21and22$`individual name`))

# create "merged_subject_id" column and add data for all variables in subject_variables list
# make sure there are no "missing" values -- change these to NA!

# na_values <- c("missing",
#                "Missing", 
#                "", 
#                " ",
#                "not collected",
#                "not applicable", 
#                "Not Applicable", 
#                "Not Collected", 
#                "N/A", 
#                "n/a",
#                "not collected",
#                "not determined",
#                "not provided",
#                "Unknown",
#                "unknown",
#                "not available",
#                "Not collected",
#                "Not applicable",
#                "Not available",
#                "na",
#                "NOT COLLECTED")
# 
# for (variable in subject_variables) {
#   for (replace_value in na_values) {
#     dataset21and22[[variable]] <- replace(dataset21and22[[variable]], dataset21and22[[variable]] == replace_value, NA)
#   }
# }

# dataset21and22$merged_subject_id <- NA

subject_variables <- c(
  "host subject id",
  "Test person",
  "submitted subject id",
  "patient_name",
  "Patient_name",
  "patient",
  "individual name",
  "individual",
  "subject_id"
)

# for (variable in subject_variables){
#   for (n_row in 1:nrow(dataset21and22)) {
#     if (is.na(dataset21and22[n_row, "merged_subject_id"])) {
#       if (!is.na(dataset21and22[[n_row, variable]])) {
#         dataset21and22[n_row, "merged_subject_id"] <- dataset21and22[[n_row, variable]]
#       }
#     }
#   }
# }

sum(!is.na(dataset21and22$merged_subject_id))
# 7159 samples with subject id

length(unique(dataset21and22$merged_subject_id))
# 2079 unique subject ids! not that bad, but i suppose it really depends on the available metadata...
```

```{r}
# create new df with subject id as row

OTU_names <- colnames(dataset[11:45])
OTU_names <- append(OTU_names, colnames(dataset[146:158]))
OTU_names <- append(OTU_names, colnames(dataset[448:454]))

# make empty df with column names from original dataset
subject_df <- data.frame(matrix(nrow = 0, ncol = ncol(dataset)))
colnames(subject_df) <- colnames(dataset)

# iterate over rows in original dataset and add rows with subject ID's one at a time
for (n_row in 1:nrow(dataset)) {
  # check that there is a value for subject id
  if (!is.na(dataset[n_row, "merged_subject_id"])) {
    # add row to subject_df if subject id isn't already in there
    if (!dataset[n_row, "merged_subject_id"] %in% as.list(subject_df$merged_subject_id)){
      subject_df <- rbind(subject_df, dataset[n_row,])
    }
    # if subject id already in df, append sample_id to sample_id column and iterate over OTU names (retain TRUE)
    else if (dataset[n_row, "merged_subject_id"] %in% as.list(subject_df$merged_subject_id)) {
      sample_ids <- append(subject_df$sample_id[subject_df$merged_subject_id == dataset[n_row, "merged_subject_id"]], dataset[n_row, "sample_id"])
      subject_df$sample_id[subject_df$merged_subject_id == dataset[n_row, "merged_subject_id"]] <- toString(sample_ids)
      # retain TRUE ("1") values for OTUs
      for (OTU in OTU_names) {
        # check that OTU = 1 in original df
        if (dataset[n_row, OTU] == 1) {
          # check that OTU = 0 in subject df
          if (subject_df[,OTU][subject_df$merged_subject_id == dataset[n_row, "merged_subject_id"]] == 0) {
            # change OTU 0 --> 1 in subject df
            subject_df[,OTU][subject_df$merged_subject_id == dataset[n_row, "merged_subject_id"]] <- 1
          }
        }
      }
    }
  }
}

```

```{r}
# recalculate total num OTUs per subject (in subject df)
# (just copy the code used to calculate total num OTUs per sample ...)

subject_df$"n_leaf-level_OTUs_11to45" <- rowSums(subject_df[11:45])
subject_df$"n_leaf-level_OTUs_146to158" <- rowSums(subject_df[146:158])
subject_df$"n_leaf-level_OTUs_448to454" <- rowSums(subject_df[448:454])

subject_df$"n_leaf-level_OTUs" <- rowSums(subject_df[457:459])

subject_df$SRM_present <- subject_df$`n_leaf-level_OTUs` > 0

```

# finding project accession for every. damn. biosample.
# yes, you heard me right
```{r}
# working df = data (copy of dataset21and22)

# # next, let's get a list of the samples that fall into this category
# samples <- as.vector(dataset21and22$sample_id)

# and now, we check if these samples have any metadata variables related to a project number
# sum(!is.na(data$ProjectAccession))
# 731 biosamples have info for project accession, so we can add these directly to the nmerged project_accession column

# welp. none of the samples have a project accession. 
# new strategy: 
# step 1: manually find project accession for sample 
# step 2: download list of samples associated with that project 
# step 3: match samples from list in step 2 ("project_samples") with samples in "samples" and add to "matched_samples" list
# step 4: add project accession to new "project_accession" column for matched samples dataset21and22
# step 5: remove matched samples from "samples" -- wash rinse repeat

# first, add all values from ProjectAccession to merged_project_accession
# data$merged_project_accession <- data$ProjectAccession

PRJDB8054 <- read_csv("Bioproject_Biosample_Lists/PRJDB8054.txt", col_names = FALSE, show_col_types = FALSE)
projects <- c(PRJNA514245, PRJNA442434, PRJNA178162, PRJEB26092, PRJDB4597, PRJDB6127, PRJDB6498, PRJDB6872, PRJDB7521, PRJDB7714, PRJDB8054)
project_accession <- "PRJDB8054"
project_samples <- as.vector(tail(projects, 1)[[1]])

matched_samples <- intersect(project_samples, samples)

# first project
# data$merged_project_accession <- ifelse(data$sample_id %in% matched_samples, project_accession, NA)

# subsequent projects
data$merged_project_accession <-
ifelse(is.na(data$merged_project_accession),
         ifelse(data$sample_id %in% matched_samples, project_accession, NA), data$merged_project_accession)

length(matched_samples)
# View(table(data$merged_project_accession))
sum(!is.na(data$merged_project_accession))

samples <- as.vector(subset(data$sample_id, is.na(data$merged_project_accession)))
samples[1]
```


```{r}
# projects with only one sample

sample <- samples[1]
proj_acc <- "PRJDB4440"

data[data$sample_id == sample, "merged_project_accession"] <- proj_acc

sum(!is.na(data$merged_project_accession))

samples <- as.vector(subset(data$sample_id, is.na(data$merged_project_accession)))
samples[1]
```


## OKAY. we are not doing this. i finally learned how to use NCBI's eUtils, so I wrote a bash script parser to get the project for every sample!!! exciting 
# so while that's running, i'm gonna run the same code on a shorter list of samples for samples that: 1) do not have a merged_subject_id and 2) have a value for age_years or one of the binned age columns
# so this code is to get that list of samples!
```{r}
# n samples that have merged_subject_id = 7,159
length(as.vector(subset(dataset21and22$sample_id, !is.na(dataset21and22$merged_subject_id))))

# n samples that have merged_subject_id and age_years = 4,405
length(as.vector(subset(dataset21and22$sample_id, !is.na(dataset21and22$merged_subject_id) & (!is.na(dataset21and22$age_years) | !is.na(dataset21and22$bin_9yrSize) | !is.na(dataset21and22$bin_5yrCutoff)))))

# n samples that have merged_subject_id but not age_years = 2,754
7159 - 4405

# n samples that have age_years, bin_9yrSize, or bin_5yrCutoff but not merged_subject_id = 4,703
length(as.vector(subset(dataset21and22$sample_id, is.na(dataset21and22$merged_subject_id) & (!is.na(dataset21and22$age_years) | !is.na(dataset21and22$bin_9yrSize) | !is.na(dataset21and22$bin_5yrCutoff)))))

samples <- as.vector(subset(dataset21and22$sample_id, is.na(dataset21and22$merged_subject_id) & (!is.na(dataset21and22$age_years) | !is.na(dataset21and22$bin_9yrSize) | !is.na(dataset21and22$bin_5yrCutoff))))

write_lines(samples, file = "biosample_accession_withAgeData_noSubjectID.txt", sep = "\n")

```


