---
title: "charity_work_for_tloyl"
output: html_document
---

# DESCRIPTION
# My task and ultimate goal is to analyze the co-occurence of 55 bacterial species within a cohort of 1000+ subjects. There's a (supposedly) convenient package in R, cooccur, that calculates the probabilistic species cooccurrence of species pairs from a simple site by species (or species by site) matrix as the input. In this case, each "site" is a subject. Starting an excel spreadsheet of the site-species matrix, this markdown file will walk you through the steps needed to apply the cooccur function.

# packages to install
```{r}
library(cooccur)
library(xlsx)
```

# read in the excel spreadsheet as a site-species matrix
```{r}
# details on the spreadsheet:
# - Row names are in column 1. Each row represents a subject. There are 1,294 subjects/sites/rows.
# - Column names are in row 1. Each column represents a species. There are 55 species.
# - Cell values are binary. Zero (0) indicates the species was NOT identified in that subject. One (1) indicates the species was identified in that subject. This is presence/absence data, not species abundance. 

subject_OTU_matrix <- read.xlsx2("subject_OTU_matrix.xlsx", row.names=1, sheetIndex = 1, colClasses = c("character", rep("numeric", 54)), endRow = 1295)
View(subject_OTU_matrix)
```

# (attempted/failed) cooccurrence analysis
```{r}
# Let's test out the cooccur function with a subset of the site-species matrix. 
subject_OTU_matrix <- subject_OTU_matrix[1:9,1:3]

subject_OTU_matrix_transposed <- t(subject_OTU_matrix)

# and now for the cooccur function...
cooccur.subjectOTU <- cooccur(mat=subject_OTU_matrix_transposed, spp_names = T)
```

# let's look at results!
```{r}
# summary(cooccur.subjectOTU)
# prob.table(cooccur.subjectOTU)
plot(cooccur.subjectOTU)
```

# making new species-subject matrix, transferring OTU-level data to to species-level (to the best of my ability lol)
```{r}
# blank subject-species matrix (species level instead of OTU level)
subject_species_matrix <- matrix(data = 0,
       nrow = nrow(subject_OTU_matrix),
       ncol = length(unique(OTU_species_table$`90% nt cluster`)))

rownames(subject_species_matrix) <- (rownames(subject_OTU_matrix))
colnames(subject_species_matrix) <- c(unique(OTU_species_table$`90% nt cluster`))

# adding data! woo! :|

# testing with mini df
# subject_species_matrix <- subject_species_matrix[1:10, 1:10]
# subject_OTU_matrix <- subject_OTU_matrix[1:10, 1:10]
# colnames(subject_species_matrix)

# for (species in colnames(subject_species_matrix)) {
#   OTUs <- (subset(OTU_species_table$OTU, OTU_species_table$species == species))
#   # OTUs <- colnames(subject_OTU_matrix)
#   for (OTU in OTUs) {
#     subject_species_matrix[,species] <- ifelse((subject_OTU_matrix[,OTU] == 1 &
#                                                        subject_species_matrix[,species] == 0),
#                                                      1,
#                                                      0)
#   }
# }

# uhh ok im just gonna iterate by row ..........

for (species in colnames(subject_species_matrix)) {
  OTUs <- (subset(OTU_species_table$OTU, OTU_species_table$`90% nt cluster` == species))
  
  for (n_row in 1:nrow(subject_species_matrix)) {
    for (OTU in OTUs) {
      if (subject_OTU_matrix[n_row, OTU] == 1) {
        subject_species_matrix[n_row, species] <- 1
      }
    }
  }
}

```

# species-level cooccurrence
```{r}
subject_species_matrix_transposed <- t(subject_species_matrix)

# subsetting the df to take out some species with <20 observations 
species_to_exclude <- c("Desulfitobacterium hafniense", "Desulfovibrio desulfuricans cluster 3", "Thermodesulfobium narugense", "unclassified Firmicute Unc01cxc", "Archaeoglobus sulfaticallidus", "Desulfovibrio desulfuricans cluster 2")
species_to_include <- setdiff(rownames(subject_species_matrix_transposed), species_to_exclude)
subject_species_matrix_transposed_subset <- subject_species_matrix_transposed[rownames(subject_species_matrix_transposed) %in% species_to_include,]
cooccur.subjectSpecies <- cooccur(mat=subject_species_matrix_transposed_subset, spp_names = T)

# summary(cooccur.subjectSpecies)
# prob.table(cooccur.subjectSpecies)

# plot(cooccur.subjectOTU)
# plot(cooccur.subjectSpecies) + 
#   theme(legend.position = c(0.93, 0.5), plot.title = element_text(size=15), legend.text=element_text(size=10))
  
```

# number of observations (sites/subjects w/ pos for each species) from subejct_species_matrix
```{r}

# species_occurrence_table <- matrix(ncol= 1, nrow = ncol(subject_species_matrix))
# rownames(species_occurrence_table) <- colnames(subject_species_matrix)
# colnames(species_occurrence_table) <- "n_observations"
# 
# 
# species_occurrence_table$"n_observations" <- ifelse()
# 


species_occurrence_table <- subject_species_matrix_transposed_subset

species_occurrence_table %>% 
  mutate(n_subjects = colSums(.))

View(species_occurrence_table)

```


