---
title: "firmicutes_tree_zoom"
output: html_document
---

#packages to load
```{r}
library(ggplot2)
library(treeio)
library(ggtree)
library(dplyr)
library(ape)
```

# prepare result tree! load, root, select query and ref leaves, remove Moorella cpy 2
```{r}
# load result tree in and root
result_tree = read.tree("/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/main/workflow/out/raxmlOutputFinal/RAxML_labelledTree_noBootstrap_rooted.newick")
rooted_result_tree <-root(result_tree, c("Moorella_thermoacetica_Y72_copy_2__MooTher5"), r = TRUE)

# only keep leaves that are in the closest ref leaves list OR have "query" in the name
closest_ref_leaves <- as.vector(otu_hits_table$Var1)
# closest ref leaves from otu_hits_table (made from the ref tree)
query_leaves <- as.vector(subset(rooted_result_tree$tip.label, grepl("QUERY", rooted_result_tree$tip.label) == TRUE))

# remove Moorella copy 2 from tree
trimmed_result_tree <- keep.tip(rooted_result_tree, c(closest_ref_leaves, query_leaves))
moorella_queries <- queryDistanceInfo$query_names[grepl("copy_2", queryDistanceInfo$closest_ref) == TRUE]

tree <- drop.tip(trimmed_result_tree, c(moorella_queries, "Moorella_thermoacetica_Y72_copy_2__MooTher5"))

```

# subset tree for Firmicutes node and visualize
```{r}
# using viewClade function
# firmicutes_node <- MRCA(tree, "Gordonibacter_pamelaeae__GrdPame5", "Polluted_aquifer_clone_LGWK16__UncS1633")
# tree_plot <- ggtree(tree) +
# #   geom_tiplab(aes(subset= label %in% closest_ref_leaves), size = 2, align = TRUE)
# viewClade(tree_plot, firmicutes_node)

# using tree_subset function
# firmicutes_subset_tree <- tree_subset(tree, "Gordonibacter_pamelaeae__GrdPame5", levels_back = 7)
firmicutes_subset_tree <- tree_subset(tree, firmicutes_node, levels_back = 0)
ggtree(firmicutes_subset_tree, layout = "circular") +
  geom_tiplab(aes(subset= label %in% closest_ref_leaves), size = 2, align = TRUE)
ggtree(firmicutes_subset_tree) +
  geom_tiplab(aes(subset= label %in% closest_ref_leaves), size = 2, align = TRUE)
gordonibacter_pamaleae_node <- MRCA(tree, "Gordonibacter_pamelaeae__GrdPame5", "Gordonibacter_pamelaeae_7-10-1-b__FP929047")
# tree_plot %>% collapse(node = 1883)

# collapse Desulfitibacter alkalitolerans surrounding node to viz half of firmicutes tree
# upperFirmNode <- MRCA(firmicutes_subset_tree, "Desulfitibacter_alkalitolerans__DfiAlka2", "QUERY___S233")
# color = label %in% closest_ref_leaves
tree_plot <- ggtree(firmicutes_subset_tree, layout = "slanted", aes(alpha = label %in% closest_ref_leaves, color = label %in% closest_ref_leaves)) +
# geom_tiplab(aes(subset= label %in% closest_ref_leaves), size = 2, align = TRUE, offset = -0.5) +
  scale_color_manual(values = c("grey", "black")) +
  scale_alpha_manual(values = c(0.2, 1)) +
  guides(color = "none", alpha = "none") +
  geom_treescale(x = 0.09, y = 300, width = 0.5, fontsize = 4, color = "darkgray") +
  
  # reference leaf label
  geom_cladelabel(node = MRCA(firmicutes_subset_tree, "Gordonibacter_pamelaeae__GrdPame5"), label = 'paste(italic("Gordonibacter pamelaeae"))', parse = T, fontsize = 4, offset = 0.0) +
  
  geom_cladelabel(node = MRCA(firmicutes_subset_tree, "Gordonibacter_pamelaeae_7-10-1-b__FP929047"), label = 'paste(italic("Gordonibacter pamelaeae"))', parse = T, fontsize = 4, offset = 0.0) +

  geom_cladelabel(node = MRCA(firmicutes_subset_tree, "Desulfitibacter_alkalitolerans__DfiAlka2"), label = 'paste(italic("Desulfitibacter alkalitolerans"))', parse = T, fontsize = 4, offset = -.01) +

  geom_cladelabel(node = MRCA(firmicutes_subset_tree, "Desulfitobacterium_hafniense_Y51__DsfHaf15"), label = 'paste(italic("Desulfitobacterium hafniense"))', parse = T, fontsize = 4, offset = 0.0) +

    geom_cladelabel(node = MRCA(firmicutes_subset_tree, "Desulfitobacterium_sp._PCE-1__DstSpe13"), label = 'paste(italic("Desulfitobacterium "), "spp.")', parse = T, fontsize = 4, offset = -0.75) +

    geom_cladelabel(node = MRCA(firmicutes_subset_tree, "Desulfosporosinus_youngiae__DfpYoun5"), label = 'paste(italic("Desulfosporosinus youngiae"))', parse = T, fontsize = 4, offset = 0) +

  geom_cladelabel(node = MRCA(firmicutes_subset_tree, "Thermodesulfobium_narugense__CP002690"), label = 'paste(italic("Thermodesulfobium narugense"))', parse = T, fontsize = 4, offset = 0)

unclassified_bacteria <- c("Anaerobic_bacterium_sk.prop8__Bv2Prop2", "Anaerobic_bacterium_EtOH8__Bv2EtOH2", "MSW_digester_clone__Unc01cxc", "Deep_sea_sediment_clone__JN798931", "Bioreactor_clone__UncS2029", "Pearl_river_estuary_clone_TopDSR14__UncS2036", "Arctic_sediment_clone_AH5_5__AtcSed67", "Polluted_aquifer_clone_LGWK16__UncS1633")
  
for (node in unclassified_bacteria) {
  tree_plot <- tree_plot +  geom_cladelabel(node = MRCA(firmicutes_subset_tree, node), label = "*", fontsize = 5, offset = 0)
}
  

# tree_plot %>%
#   collapse(node = upperFirmNode) +
#   geom_point2(aes(subset=(node==upperFirmNode)), shape=23, size=2, fill='blue') 

# tree_plot_allLabels <- ggtree(firmicutes_subset_tree, layout = "circular") +
# geom_tiplab(size = 2, align = TRUE)
# tree_plot_allLabels %>%
#   collapse(node = upperFirmNode) +
#   geom_point2(aes(subset=(node==upperFirmNode)), shape=23, size=3, fill='red')

# viewClade(tree_plot, upperFirmNode)

# add metadata on metabolism
# tree_plot %<+%
#   otu_all_table +
#   geom_point2(aes(color = updated_freq, subset = Freq > 0))
# ggsave("firmicutes_tree_plot.png", tree_plot, height = 4000, units = "px")

tree_plot

```


```{r}
tree_plot %<+%
otu_all_table +
geom_point2(aes(color = updated_freq, subset = Freq > 0))

# tree_plot
```

# subset desulfovibrionaceae node and zoom in to visualize
```{r}
dsl_node <- MRCA(tree, "uncultured_bacterium__EF645666", "Desulfovibrio_desulfuricans__DslDes24")

dsl_subset_tree <- tree_subset(tree, dsl_node, levels_back = 0)

upperDisNode <- MRCA(dsl_subset_tree, "Desulfovibrio_piger__DslPige3", "QUERY___S606")

dsl_tree_plot <- ggtree(dsl_subset_tree, layout = "slanted", aes(color = label %in% closest_ref_leaves, alpha = label %in% closest_ref_leaves)) +
# geom_tiplab(aes(subset= label %in% closest_ref_leaves), size = 4, align = F) +
  # geom_tiplab() +
  scale_color_manual(values = c("grey", "red")) +
  scale_alpha_manual(values = c(0.5, 1)) +
  guides(color = "none", alpha = "none")

#   geom_treescale(x = 0.09, y = 300, width = 0.1, fontsize = 4, color = "darkgray") +
#   
# # labels
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_piger__DslPige3"), label = 'paste(italic("Desulfovibrio piger"))', parse = T, fontsize = 4, offset = 0, color = "red") +
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_piger__DslPige2"), label = 'paste(italic("Desulfovibrio piger"))', parse = T, fontsize = 4, offset = 0, color = "red") +
# 
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_desulfuricans__DQ092635"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 4, offset = 0, color = "red") +
# 
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_desulfuricans_subsp._desulfuricans_str._ATCC_27774__CP001358"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 4, offset = 0.2, color = "red") +
# 
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_simplex__DslSimpl"), label = 'paste(italic("Desulfovibrio simplex"))', parse = T, fontsize = 4, offset = -0.7, color = "red") +
# 
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_sp._3_1_syn3__DsvSp230"), label = 'paste(italic("Desulfovibrio "), "spp.")', parse = T, fontsize = 4, offset = -0.7, color = "red") +
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_sp._Dsv1__DsvSp233"), label = 'paste(italic("Desulfovibrio "), "spp.")', parse = T, fontsize = 4, offset = 0, color = "red") +
# 
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Candidatus_Desulfovibrio_trichonymphae"), label = 'paste("Candidatus ", italic("Desulfovibrio trichonymphae"))', parse = T, fontsize = 4, offset = -0.7, color = "red") +
# 
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Bilophila_sp._4_1_30__BilSpec3"), label = 'paste(italic("Bilophila "), "spp.")', parse = T, fontsize = 4, offset = 0, color = "red") +
# 
#     geom_cladelabel(node = MRCA(dsl_subset_tree, "Bilophila_wadsworthia__BilWads6"), label = 'paste(italic("Bilophila wadsworthia"))', parse = T, fontsize = 4, offset = 0, color = "red") +
# 
#   geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_vulgaris__DQ826729"), label = 'paste(italic("Desulfovibrio vulgaris"))', parse = T, fontsize = 4, offset = 0, color = "red") +
# 
#     geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_desulfuricans__DslDes16_DsvAlas8"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 4, offset = 0.5, color = "red") +
# 
#     geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_carbinoliphilus__JQ031012"), label = 'paste(italic("Desulfovibrio carbinoliphilus"))', parse = T, fontsize = 4, offset = 0, color = "red") +
# 
#     geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_desulfuricans__DslDes24"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 4, offset = -0.6, color = "red")
# 
# unclassified_bacteria <- c("uncultured_bacterium__EF645666", "uncultured_bacterium__EF645667", "uncultured_bacterium__EF645664", "uncultured_bacterium__EF645665", "Bioreactor_clone__UncS2028", "Termite_protist_endosymbiont_community_metagenome__entry118", "Polluted_aquifer_clone__UncS1623")

# for (node in unclassified_bacteria) {
#   dsl_tree_plot <- dsl_tree_plot +  geom_cladelabel(node = MRCA(dsl_subset_tree, node), label = "*", fontsize = 5, offset = 0, color = "red")
# }

# ggsave("Tree_Plots/desulfovibrionaceae_tree_plot.png", dsl_tree_plot, height = 3000, units = "px")
```

# lower desulfovibrionaceae branches
```{r}
upperDisNode <- MRCA(dsl_subset_tree, "Desulfovibrio_piger__DslPige3", "Bilophila_wadsworthia__BilWads6")

dsl_tree_plot <- ggtree(dsl_subset_tree, layout = "slanted", aes(color = label %in% closest_ref_leaves, alpha = label %in% closest_ref_leaves)) +
# geom_tiplab(aes(subset= label %in% closest_ref_leaves), size = 4, align = F) +
  # geom_tiplab() +
  scale_color_manual(values = c("grey", "darkorange1")) +
  scale_alpha_manual(values = c(0.5, 1)) +
  guides(color = "none", alpha = "none")

dsl_tree_plot_lower <- dsl_tree_plot %>%
  collapse(node = upperDisNode) +
  geom_point2(aes(subset=(node==upperDisNode)), shape=23, size=3, fill='darkorange1', color = 'darkorange1', alpha = 1) +
  
  # labels
    geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_vulgaris__DQ826729"), label = 'paste(italic("Desulfovibrio vulgaris"))', parse = T, fontsize = 3, offset = 0, color = "darkorange1") +
    geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_desulfuricans__DslDes16_DsvAlas8"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 3, offset = 0, color = "darkorange1") +
  geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_carbinoliphilus__JQ031012"), label = 'paste(italic("Desulfovibrio carbinoliphilus"))', parse = T, fontsize = 3, offset = 0, color = "darkorange1") +
  geom_cladelabel(node = MRCA(dsl_subset_tree, "Desulfovibrio_desulfuricans__DslDes24"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 3, offset = -0, color = "darkorange1") +
  
    geom_treescale(x = 0.06, y = 50, width = 0.1, fontsize = 3, color = "darkgray")

  
unclassified_bacteria <- c("Polluted_aquifer_clone__UncS1623")

for (node in unclassified_bacteria) {
  dsl_tree_plot_lower <- dsl_tree_plot_lower +  
    geom_cladelabel(node = MRCA(dsl_subset_tree, node), label = "*", fontsize = 5, offset = 0, color = "darkorange1")
}

dsl_tree_plot_lower
# ggsave("Tree_Plots/desulfovibrionaceae_tree_plot_lowerHalf.png", dsl_tree_plot_lower, height = 2000, units = "px")
```

# upper desulfovibrionaceae branches
```{r}
upperDisNode <- MRCA(dsl_subset_tree, "Desulfovibrio_piger__DslPige3", "Bilophila_wadsworthia__BilWads6")

dsl_subset_tree_upper <- tree_subset(dsl_subset_tree, upperDisNode, levels_back = 0)

dsl_tree_plot_upper <- ggtree(dsl_subset_tree_upper, layout = "slanted", aes(color = label %in% closest_ref_leaves, alpha = label %in% closest_ref_leaves)) +
# geom_tiplab(aes(subset= label %in% closest_ref_leaves), size = 2, align = T) +
  # geom_tiplab() +
  scale_color_manual(values = c("grey", "darkorange1")) +
  scale_alpha_manual(values = c(0.3, 1)) +
  guides(color = "none", alpha = "none") +
  geom_treescale(x = 0.02, y = 1200, width = 0.1, fontsize = 3, color = "darkgray") +
  ggplot2::ylim(-0.1, 1300) +

# labels
  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Desulfovibrio_piger__DslPige3"), label = 'paste(italic("Desulfovibrio piger"))', parse = T, fontsize = 2, offset = 0, color = "darkorange1") +
  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Desulfovibrio_piger__DslPige2"), label = 'paste(italic("Desulfovibrio piger"))', parse = T, fontsize = 2, offset = 0, color = "darkorange1") +

  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Desulfovibrio_desulfuricans__DQ092635"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 2, offset = 0.01, color = "darkorange1") +

  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Desulfovibrio_desulfuricans_subsp._desulfuricans_str._ATCC_27774__CP001358"), label = 'paste(italic("Desulfovibrio desulfuricans"))', parse = T, fontsize = 2, offset = 0.03, color = "darkorange1") +

  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Desulfovibrio_simplex__DslSimpl"), label = 'paste(italic("Desulfovibrio simplex"))', parse = T, fontsize = 2, offset = -0.15, color = "darkorange1") +

  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Desulfovibrio_sp._3_1_syn3__DsvSp230"), label = 'paste(italic("Desulfovibrio "), "spp.")', parse = T, fontsize = 2, offset = -0.15, color = "darkorange1") +
  # geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Desulfovibrio_sp._Dsv1__DsvSp233"), label = 'paste(italic("Desulfovibrio "), "spp.")', parse = T, fontsize = 3, offset = 0, color = "darkorange1") +

  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Candidatus_Desulfovibrio_trichonymphae"), label = 'paste("Candidatus ", italic("Desulfovibrio trichonymphae"))', parse = T, fontsize = 2, offset = 0.1, color = "darkorange1") +

  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Bilophila_sp._4_1_30__BilSpec3"), label = 'paste(italic("Bilophila "), "spp.")', parse = T, fontsize = 2, offset = 0, color = "darkorange1") +

    geom_cladelabel(node = MRCA(dsl_subset_tree_upper, "Bilophila_wadsworthia__BilWads6"), label = 'paste(italic("Bilophila wadsworthia"))', parse = T, fontsize = 2, offset = 0, color = "darkorange1")

unclassified_bacteria <- c("uncultured_bacterium__EF645666", "uncultured_bacterium__EF645667", "uncultured_bacterium__EF645664", "uncultured_bacterium__EF645665", "Bioreactor_clone__UncS2028", "Termite_protist_endosymbiont_community_metagenome__entry118")

for (node in unclassified_bacteria) {
  dsl_tree_plot_upper <- dsl_tree_plot_upper +  geom_cladelabel(node = MRCA(dsl_subset_tree_upper, node), label = "*", fontsize = 5, offset = -0.001, color = "darkorange1")
}

dsl_tree_plot_upper

ggsave("Tree_Plots/desulfovibrionaceae_tree_plot_upperHalf.png", dsl_tree_plot_upper, height = 2000, units = "px")

```


# desulfovibrionaceae tree, annotated with tip shapes instead of labels
```{r}
# dsl_node <- MRCA(tree, "uncultured_bacterium__EF645666", "Desulfovibrio_desulfuricans__DslDes24")

# dsl_subset_tree <- tree_subset(tree, dsl_node, levels_back = 0)

dsl_tree_plot_points <- ggtree(dsl_subset_tree, layout = "slanted", alpha = 0.3, color = "gray")

dsl_tree_plot_points %<+% desulfovibrionaceae_species_metadataTable +
   geom_tippoint(aes(subset = label %in% closest_ref_leaves, color = species))
# ggsave("Tree_Plots/desulfovibrionaceae_tree_plot_tipPoints.png", dsl_tree_plot_points , height = 4000, units = "px")

```

# adding metabolism info per ref leaf to metadata (otu_all_table)
```{r}
df <- otu_all_table

sulfite_reduc

```
