---
title: "metadata_significance_analysis_bySubject"
output: html_document
---

# Sex vs. SRM Presence
```{r}

# Pearson's Chi Square Test

# contigency table of SRM presence vs. sex
table(hgm_subject_df$sex_all, hgm_subject_df$SRM_present)
chistest_sex <- chisq.test(hgm_subject_df$sex_all, hgm_subject_df$SRM_present)
chistest_sex

# Fisher's exact test
contigency_table <- table(hgm_subject_df$sex_all, hgm_subject_df$SRM_present)
fishertest <- fisher.test(contigency_table)
fishertest$p.value

# Expected values for all cells > 5 count
# Assumptions for Chi-Square are met

# p-value for 1 df = 1
# Fail to reject the null hypothesis at the 0.05 significance level. No significant difference in SRM presence vs. sex.

```

# Age vs. SRM Presence
```{r}
# nonparametric test to compare mean age (years) of SRM present vs. absent groups
# distribution of ages is not normal, so must use non-parametric test
# Signifance test: Mann-Whitney U / Wilcoxon-rank-sum Test
# bc not normal distribution, can only compare means and not medians

# wilcox.test(data = hgm_subject_df, age_years ~ SRM_present, alternative = c("less"))

# p-value < 2.2e-16
# Reject the null hypothesis at the 0.05 significance level!

# check correlation adults (> 5 years)?

# ggplot() +
# geom_boxplot(data = subset(hgm_subject_df, bin_5yrCutoff == "[5,110)"), aes(x = age_years, y = SRM_present))
pink = brewer.pal(name = "RdPu", 2)[2]
grey = brewer.pal(name = "Greys", 4)[3]

plot <- ggplot(data = hgm_subject_df, aes(x = age_years, y = SRM_present, color = SRM_present))+
  geom_violin() +
  geom_boxplot(width = 0.1) +
  # ggtitle("Age vs. SRM Presence") +
  scale_color_manual(values = c(pink, grey)) +
  # xlab("Age (years)") +
  # ylab("dsrAB Detected") +
  guides(color = "none") +
  theme_bw() +
  scale_x_continuous(breaks = seq(0,90, 10)) +
  geom_signif(comparisons = list(c("FALSE", "TRUE")), test = "wilcox.test", map_signif_level = T, color = "black") +
  theme(axis.text.y = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(size = 11))

ggsave(plot = plot, file = "age_v_SRMpresence.jpeg", width = 6, height = 4)

# ggplot(data = subset(hgm_subject_df, !is.na(hgm_subject_df$bin_5yrCutoff)), aes(x = bin_5yrCutoff, fill = SRM_present)) +
# geom_bar(position = "fill", alpha = 0.8) +
# xlab("Age Bin (years)") +
# ggtitle("Proportion of SRM-Present Subjects by Age, Binned at 5 y.o.") +
#   ylab("Proportion") +
#   theme_bw() +
#   geom_text(aes(label = signif((..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)]), digits = 3)), stat = "count", position = "fill", vjust = 1.6, size = 4) +
#   scale_fill_manual(values = c("cornflowerblue", "darkorange"), name = "dsrAB Detected") +
#   guides(alpha = "none")
# geom_signif(comparisons = list(c("[0,5)", "[5,110)")), test = "fisher.test")
  
# ggplot(data = subset(hgm_subject_df, !is.na(hgm_subject_df$bin_5yrCutoff)), aes(x = bin_5yrCutoff, fill = SRM_present)) +
# geom_bar(alpha = 0.8, position = ) +
# xlab("Age Bin (years)") +
# ggtitle("Proportion of SRM-Present Subjects by Age, Binned at 5 y.o.") +
#   ylab("Proportion") +
#   theme_bw() +
#   # geom_text(aes(label = signif((..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)]), digits = 3)), stat = "count", position = "fill", vjust = 1.6, size = 4) +
#   scale_fill_manual(values = c("cornflowerblue", "darkorange"), name = "dsrAB Detected") +
#   guides(alpha = "none")
  
```
# comparison of proportion of SRM present vs absent subjects in high vs. low age groups
```{r}
# use Fischer test
# it's the nonparametric equivalent of the two-proportion Z test

contigency_table <- table(hgm_subject_df$bin_5yrCutoff, hgm_subject_df$SRM_present)
fisher.test(contigency_table)$p.value

# now add significance layer to plot
# data <- subset(hgm_subject_df, !is.na(hgm_subject_df$bin_5yrCutoff))

# ggplot(data = data, aes(x = bin_5yrCutoff, fill = SRM_present)) +
# geom_bar(position = "fill", alpha = 0.8) +
# xlab("Age Bin (years)") +
# ggtitle("Proportion of SRM-Present Subjects by Age, Binned at 5 y.o.") +
#   ylab("Proportion") +
#   theme_bw() +
#   geom_text(aes(label = signif((..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)]), digits = 3)), stat = "count", position = "fill", vjust = 1.6, size = 4) +
#   scale_fill_manual(values = c("cornflowerblue", "darkorange"), name = "dsrAB Detected") +
#   guides(alpha = "none") +
#   geom_signif(comparisons = list(c("[0,5)", "[5,110)")), test = "fisher.test", aes(y = stat(count/sum(count))))

data %>%
    group_by(x = bin_5yrCutoff) %>%
    count(y = SRM_present) %>%
    # mutate(Freq = n / sum(n)) %>%
    ggplot() +
    aes(x, n, fill = y) +
    geom_col(position = position_dodge(0.9)) +
    geom_signif(
        comparisons = list(c("[0,5)", "[5,110)")),
        map_signif_level = TRUE
    )

```

# Country vs. SRM Presence
```{r}

# subset by countries with SRM
ggplot()+
  geom_bar(data = subset(hgm_subject_df, country %in% c("USA", "Sweden", "Russia", "Luxembourg", "Estonia", "Denmark", "China", "Austria", "Germany")), aes(y = country, fill = SRM_present), position = "fill") +
  xlab("proportion") +
  ggtitle("SRM Presence vs. Country (countries w/ > 5 subjects w/ SRM)")

# subset by countries with > 50 subjects
ggplot()+
  geom_bar(data = subset(hgm_subject_df, country %in% c("USA", "China", "Sweden", "Fiji", "Denmark", "Spain", "Russia", "Estonia", "Finland", "Austria")), aes(y = country, fill = SRM_present), position = "fill") +
  xlab("proportion") +
  ggtitle("SRM Presence vs. Country (countries w/ > 50 subjects total)")

# What question do we want to test for significance in the data?
# Ex: Does the US have a higher rate of SRM-present guts than other countries around the world?

```

```{r}
ggplot(data = subset(hgm_subject_df,!is.na(age_years) & country == "Denmark"), aes(x = age_years)) +
  geom_boxplot(show.legend = FALSE) +
  xlab("Age in years") +
  ggtitle("Age distribution by country for subjects with SRM") +
  # facet_wrap(~SRM_present) +
  scale_x_continuous(breaks = seq(0,80, 10))
```

# proportion test for 5 y.o. bin category by SRM presence
```{r}
# using infer packages
# library(infer)

# prop_test(hgm_subject_df, bin_5yrCutoff ~ SRM_present)  

# quantiles for hgm_subject_fd for age distribution in SRM present vs absent groups
# percentage of subjects under 5 without SRM
(nrow(subset(hgm_subject_df, age_years < 5 & SRM_present == F)) / nrow(subset(hgm_subject_df, !is.na(age_years) & SRM_present == F))) * 100

```

# point biserial correlation test for age (continuous) vs. SRM presence (binary)
```{r}
age_df <- subset(hgm_subject_df, !is.na(age_years))

age_df$SRM_present_binary <- ifelse(age_df$SRM_present == TRUE,
                                    1, 
                                    0
                                    )

age_cor <- cor.test(age_df$age_years, age_df$SRM_present_binary)
age_cor$p.value
age_cor

# for sex
# jk this is not point-biserial, we gotta use a fisher's exact test

# sex_df <- subset(hgm_subject_df, !is.na(sex_all))
# 
# sex_df$SRM_present_binary <- ifelse(sex_df$SRM_present == TRUE,
#                                     1, 
#                                     0
#                                     )
# 
# age_cor <- cor.test(age_df$age_years, age_df$SRM_present_binary)
# age_cor$p.value
# age_cor
```


# binning at 3 years
```{r}
# subject_df <- subject_df %>%
#   mutate(bin_3yrCutoff = cut(age_years, breaks = c(0,3,110), right = FALSE))

ggplot(data = subset(subject_df, !is.na(subject_df$bin_3yrCutoff)), aes(x = bin_3yrCutoff, fill = SRM_present)) +
geom_bar(position = "fill", alpha = 0.8) +
xlab("Age Bin (years)") +
ggtitle("Proportion of SRM-Present Subjects by Age, Binned at 5 y.o.") +
  ylab("Proportion") +
  theme_bw() +
  geom_text(aes(label = signif((..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)]), digits = 3)), stat = "count", position = "fill", vjust = 1.6, size = 4) +
  scale_fill_manual(values = c("cornflowerblue", "darkorange"), name = "dsrAB Detected") +
  guides(alpha = "none")

```

# sex
```{r}

plot <- ggplot(data = subset(hgm_subject_df, !is.na(hgm_subject_df$sex_all)), aes(x = sex_all, fill = SRM_present)) +
geom_bar(position = "fill", alpha = 0.7, width = 0.4) +
  theme_bw() +
  geom_text(aes(label = signif((..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)]), digits = 3)), stat = "count", position = "fill", vjust = 1.6, size = 4) +
  scale_fill_manual(values = c(pink, grey), name = "dsrAB Detected") +
  guides(alpha = "none", fill = "none") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), title = element_blank(), axis.text = element_text(size = 11))

ggsave(plot = plot, file = "sex_v_SRMpresence.jpeg", width = 5, height = 5)

```

# FIGURE 2: use grid function (grid arrange?) to make one panel output for all of figure 2
# age = c, sex = B, get rid of A, change age panel to 3 yr bins