---
title: "tree_visualization"
output: html_document
---

#packages to load
```{r}
library(ggplot2)
library(treeio)
library(ggtree)
library(dplyr)
library(ape)
library(phytools)
```


#root tree
```{r}
ref_tree = read.tree("/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/main/workflow/out/raxmlOutputFinal/RAxML_originalLabelledTree_noBootstrap.newick")

out_group = as.vector(archaeaRootsetIDs$X1)

rooted_ref_tree <-root(ref_tree, c("Moorella_thermoacetica_Y72_copy_2__MooTher5"), r = TRUE)
```

```{r}
# library(ggplot2)
# library(ggtree)
# library(dplyr)
# 
result_tree_f = read.tree("/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/dsrAB_Biosample_Metadata_Analysis/combine_IDs_forHits/RAxML_labelledTree_noBootstrap_rooted.newick")
rooted_ref_tree_f = read.tree("/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/main/workflow/out/raxmlOutputFinal/RAxML_originalLabelledTree_noBootstrap.newick")

# otu_hits_table <- as.data.frame(table(subset(closestRefInfo_allScoreThresholdHits_singleRefOnly$closest_single_ref, closestRefInfo_allScoreThresholdHits_singleRefOnly$distance < 2.5)))
# otu_hits_table$logFreq <- log2(otu_hits_table$Freq + 1)

hit_otus <- otu_hits_table$Var1
nonhit_otus <- setdiff(rooted_ref_tree$tip.label, hit_otus)

tobind_df <- data.frame(Var1 = nonhit_otus, Freq = 0, updated_freq = 0)

otu_all_table <- rbind(otu_hits_table, tobind_df)

otu_all_table$logFreq <- log2(otu_all_table$updated_freq + 1)

# rooted_ref_tree <- ggtree(rooted_ref_tree_f, layout = "circular")

# rooted_ref_tree %<+%
#   otu_hits_table +
#   aes(color = Freq) +
#   scale_color_gradient(low = "blue", high = "orange")

```

# heatmap of entire tree (circular)
```{r}

# rooted_ref_tree <- ggtree(rooted_ref_tree_f, layout = "circular") + geom_tiplab(size = 0)
# 
# rooted_ref_tree %<+%
#   otu_hits_table +
#   aes(color = Freq) +
#   scale_color_gradient(low = "blue", high = "orange")

hits_only_tree <- keep.tip(rooted_ref_tree, c(hits_list, ox_bact_node, nitrospirae, red_arch_node))
hits_only_tree <- ggtree(hits_only_tree) + xlim_tree(15)

```

# create heatmap by clade
```{r}

# gheatmap(rtree, data= otu_subj_mat, offset = -0.35, width = 0.5, legend_title = "Number of Hits \n (log plus one)", colnames = FALSE, low = "blue", high = "orange", color = "transparent") +
#   labs(fontsize = 4) +
#   theme(legend.position = c(1, 0.25), legend.key.size = unit(0.45, "cm"))

desulfovibrionaceae_node = MRCA(rooted_ref_tree, c('Mariager_Fjord_clone_MAF65G__Af2Fjo18', "Bioreactor_clone__UncS2021"))
viewClade(rooted_ref_tree, desulfovibrionaceae_node)

firmicutes_node = MRCA(rooted_ref_tree, 'Bioreactor_clone__UncS2026', "Black_Sea_sediment_clone_BSI-11__BlackS15")
viewClade(rooted_ref_tree, firmicutes_node)
```

# nodes
```{r}
hits_only_tree <- keep.tip(rooted_ref_tree, c(hits_list, ox_bact_node, nitrospirae, red_arch_node))

ox_bact_node <- intersect(as.vector(subset(taxa_table$OTU, taxa_table$dsrab_type == "Oxidative bacterial type DsrAB")), rooted_ref_tree$tip.label)

nitrospirae <- intersect(as.vector(subset(taxa_table$OTU, taxa_table$supercluster == "Nitrospirae supercluster")), rooted_ref_tree$tip.label)

red_arch_node <- intersect(as.vector(subset(taxa_table$OTU, taxa_table$dsrab_type == "Reductive archaeal type DsrAB")), rooted_ref_tree$tip.label)

bilophila_node <- as.vector(subset(taxa_table$OTU, grepl("Bilophila", taxa_table$OTU) == TRUE))

dpiger_node <- as.vector(subset(taxa_table$OTU, grepl("Desulfovibrio_piger", taxa_table$OTU) == TRUE))

dalkalitolerans_node <- as.vector(subset(taxa_table$OTU, grepl("Desulfitibacter_alkalitolerans", taxa_table$OTU) == TRUE))

firmicutes_node <- intersect(as.vector(subset(taxa_table$OTU, taxa_table$supercluster == "FACA group sensu lato")), hits_only_tree$tip.label)

desulfovibrionaceae_node <- intersect(as.vector(subset(taxa_table$OTU, taxa_table$level3_taxon == "Desulfovibrionaceae")), hits_only_tree$tip.label)

gpam_node <- intersect(as.vector(subset(taxa_table$OTU, taxa_table$level3_taxon == "Gordonibacter")), hits_only_tree$tip.label)

dvulgaris_node <- intersect(as.vector(subset(taxa_table$OTU, grepl("Desulfovibrio_vulgaris", taxa_table$OTU) == TRUE)), hits_only_tree$tip.label)

dcarboniphilus_node <- intersect(as.vector(subset(taxa_table$OTU, grepl("Desulfovibrio_carbinoliphilus", taxa_table$OTU) == TRUE)), hits_only_tree$tip.label)

tnarugense_node <- intersect(as.vector(subset(taxa_table$OTU, grepl("Thermodesulfobium", taxa_table$OTU) == TRUE)), hits_only_tree$tip.label)

archaeoglobus_node <- intersect(as.vector(subset(taxa_table$OTU, grepl("Archaeoglobus", taxa_table$OTU) == TRUE)), hits_only_tree$tip.label)

desulfitobacterium_node <- intersect(as.vector(subset(taxa_table$OTU, grepl("Desulfitobacterium", taxa_table$OTU) == TRUE)), hits_only_tree$tip.label)

```

# tree with labels for all hit ref leaves (for reference)
```{r}
hits_only_tree <- keep.tip(rooted_ref_tree, c(hits_list, ox_bact_node, nitrospirae, red_arch_node))
hits_only_tree <- ggtree(hits_only_tree) + xlim_tree(15) + geom_tiplab(size = 2, offset = 0.3)

tree <- hits_only_tree %>%
  # collapse and label oxidative bacterial node
  collapse(node = MRCA(hits_only_tree, ox_bact_node)) %>%
  collapse(node = MRCA(hits_only_tree, red_arch_node)) %>%
  collapse(node = MRCA(hits_only_tree, nitrospirae)) +
  
  # outgroup labels
  
  geom_point2(aes(subset=(node == MRCA(hits_only_tree, ox_bact_node))), shape = 21, fill = "gray", color = "gray") +
  geom_cladelabel(node = MRCA(hits_only_tree, ox_bact_node), label = "oxidative bacterial type dsrAB", fontsize = 2.5, color = "gray", offset = 0.3) +
  
  geom_point2(aes(subset=(node == MRCA(hits_only_tree, nitrospirae))), shape = 21, fill = "gray", color = "gray") +
  geom_cladelabel(node = MRCA(hits_only_tree, nitrospirae), label = "nitrospirae supercluster", fontsize = 2.5, color = "gray", offset = 0.3) +
  
  geom_point2(aes(subset=(node == MRCA(hits_only_tree, red_arch_node))), shape = 21, fill = "gray", color = "gray") +
  geom_cladelabel(node = MRCA(hits_only_tree, red_arch_node), label = "reductive archaeal type dsrAB", fontsize = 2.5, color = "gray", offset = 0.3) +
  
  # group labels
  
  geom_cladelabel(node = MRCA(hits_only_tree, firmicutes_node), label = "Firmicutes group", fontsize = 4, offset = 5.5, alpha = 0.5) +
  
  geom_cladelabel(node = MRCA(hits_only_tree, desulfovibrionaceae_node), label = "Desulfovibrionaceae", fontsize = 4, offset = 4.5, alpha = 0.5)

tree %<+%
  otu_all_table +
  geom_point2(aes(color = logFreq, subset = updated_freq > 0)) +
  scale_color_gradient(low = "blue", high = "orange") +
  # guides(color = "none", size = "none") +
  theme(legend.position = "left")

```

# make it pretty
```{r}

# rooted_ref_tree = ape::read.tree("/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/main/workflow/out/raxmlOutputFinal/RAxML_originalLabelledTree_noBootstrap.newick")
# 
# # # only keep tips that have hits
hits_list <- as.vector(otu_hits_table$Var1)

hits_only_tree <- keep.tip(rooted_ref_tree, c(hits_list, ox_bact_node, nitrospirae, red_arch_node))

# hits_only_tree <- ggtree(hits_only_tree) + geom_tiplab(aes(subset=(label %in% otus_to_label)), size = 2) + xlim_tree(15)
# hits_only_tree <- ggtree(hits_only_tree) + geom_tiplab(aes(subset=(grepl("Desulfovibrio", label)) == TRUE), size = 2) + xlim_tree(15)

hits_only_tree <- ggtree(hits_only_tree, color = "dark gray") + xlim_tree(10)


tree <- hits_only_tree %>%
  # collapse and label oxidative bacterial node
  collapse(node = MRCA(hits_only_tree, ox_bact_node)) %>%
  collapse(node = MRCA(hits_only_tree, red_arch_node)) %>%
  collapse(node = MRCA(hits_only_tree, nitrospirae)) +

  # outgroup labels

  geom_point2(aes(subset=(node == MRCA(hits_only_tree, ox_bact_node))), shape = 21, fill = "dark gray", color = "dark gray") +
  geom_cladelabel(node = MRCA(hits_only_tree, ox_bact_node), label = "oxidative bacterial type dsrAB", fontsize = 2.5, color = "dark gray") +

  geom_point2(aes(subset=(node == MRCA(hits_only_tree, nitrospirae))), shape = 21, fill = "dark gray", color = "dark gray") +
  geom_cladelabel(node = MRCA(hits_only_tree, nitrospirae), label = "nitrospirae supercluster", fontsize = 2.5, color = "dark gray") +

  geom_point2(aes(subset=(node == MRCA(hits_only_tree, red_arch_node))), shape = 21, fill = "dark gray", color = "dark gray") +
  geom_cladelabel(node = MRCA(hits_only_tree, red_arch_node), label = "reductive archaeal type dsrAB", fontsize = 2.5, color = "dark gray") +

  # geom_point2(aes(subset=(node == MRCA(hits_only_tree, "Moorella_thermoacetica_Y72_copy_2__MooTher5"))), shape = 21, fill = "dark gray", color = "dark gray") +
  # geom_cladelabel(node = MRCA(hits_only_tree, "Moorella_thermoacetica_Y72_copy_2__MooTher5"), label = "Moorella thermoacetica, dsrAB copy 2", fontsize = 2.5, offset.text = 0.3, color = "dark gray") +

  # species labels

  geom_cladelabel(node = MRCA(hits_only_tree, bilophila_node), label = 'paste(italic("Bilophila"), " spp.* 59.7%")', parse = T, fontsize = 2.5, color = "black") +

  geom_cladelabel(node = MRCA(hits_only_tree, dpiger_node), label = 'paste(italic("Desulfovibrio piger"), "* 16.9%")', parse = T, fontsize = 2.5, color = "black") +

  geom_cladelabel(node = MRCA(hits_only_tree, dalkalitolerans_node), label = 'paste(italic("Desulfitibacter alkalitolerans"), " 11.8%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, gpam_node), label = 'paste(italic("Gordonibacter pamelaeae"), "* 6.5%")', parse = T, fontsize = 2.5, color = "black") +
  geom_cladelabel(node = MRCA(hits_only_tree, dvulgaris_node), label = 'paste(italic("Desulfovibrio vulgaris"), "* 4.0%")', parse = T, fontsize = 2.5, color = "black") +

  geom_cladelabel(node = MRCA(hits_only_tree, dcarboniphilus_node), label = 'paste(italic("Desulfovibrio carboniphilus"), " <1%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, tnarugense_node), label = 'paste(italic("Thermodesulfobium narugense"), " 1.2%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, archaeoglobus_node), label = 'paste(italic("Archaeoglobus sulfaticallidus"), " 1.3%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, desulfitobacterium_node), label = 'paste(italic("Desulfitobacterium hafniense"), " <1%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, "Desulfovibrio_simplex__DslSimpl"), label = 'paste(italic("Desulfovibrio simplex"), " <1%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, "Candidatus_Desulfovibrio_trichonymphae"), label = 'paste(italic("Candidatus"), " Desulfovibrio trichonymphae <1%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, "Desulfosporosinus_youngiae__DfpYoun5"), label = 'paste(italic("Desulfosporosinus youngiae"), " <1%")', parse = T, fontsize = 2.5) +

  geom_cladelabel(node = MRCA(hits_only_tree, c("Desulfovibrio_sp._6_1_46AFAA__DsvSp231", "Desulfovibrio_sp._3_1_syn3__DsvSp230", "Desulfovibrio_sp._Dsv1__DsvSp233")), label = 'paste(italic("Desulfovibrio"), " spp. 3.8%")', parse = T, fontsize = 2.5) +

  # Desulfovibrio desulfuricans labels
  geom_cladelabel(node = MRCA(hits_only_tree, "Desulfovibrio_desulfuricans__DQ092635"), label = 'paste(italic("Desulfovibrio desulfuricans*"), " <1%")', parse = T, fontsize = 2.5, color = "black") +
  geom_cladelabel(node = MRCA(hits_only_tree, c("Desulfovibrio_desulfuricans__DslDes17", "Desulfovibrio_desulfuricans_subsp._desulfuricans_str._ATCC_27774__CP001358")), label = 'paste(italic("Desulfovibrio desulfuricans*"), " <1%")', parse = T, fontsize = 2.5, color = "black") +
  geom_cladelabel(node = MRCA(hits_only_tree, "Desulfovibrio_desulfuricans__DslDes16_DsvAlas8"), label = 'paste(italic("Desulfovibrio desulfuricans*"), " <1%")', parse = T, fontsize = 2.5, color = "black") +
  geom_cladelabel(node = MRCA(hits_only_tree, "Desulfovibrio_desulfuricans__DslDes24"), label = 'paste(italic("Desulfovibrio desulfuricans*"), " <1%")', parse = T, fontsize = 2.5, color = "black") +

  # group labels

  # geom_cladelabel(node = MRCA(hits_only_tree, firmicutes_node), label = "Firmicutes group s.l. ", fontsize = 2.5, offset = 4, horizontal = FALSE) +
  #
  # geom_cladelabel(node = MRCA(hits_only_tree, desulfovibrionaceae_node), label = "Desulfovibrionaceae", fontsize = 2.5, offset = 3, horizontal = FALSE)

    geom_cladelabel(node = MRCA(hits_only_tree, firmicutes_node), label = "", fontsize = 2.5, offset = 4, horizontal = FALSE) +

  geom_cladelabel(node = MRCA(hits_only_tree, desulfovibrionaceae_node), label = "", fontsize = 2.5, offset = 3, horizontal = FALSE)


# tree %<+%
#   otu_all_table +
#   geom_point2(aes(color = logFreq, subset = Freq > 0)) +
#   scale_color_gradient(low = "orange", high = "blue") +
#   # guides(color = "none") +
#   theme(legend.position = "left")


```

# aesthetic mappings for the tree
```{r}
tree <-
  tree %<+%
  otu_all_table +
  geom_point2(aes(color = logFreq, subset = updated_freq > 0)) +
  scale_color_gradient(low = "orange", high = "blue", name = "") +
  # guides(color = "none") +
  theme(legend.position = c(0.09, 0.85), legend.key.size = unit(0.5, "cm"),legend.title = element_text(size=6)) +
  geom_treescale(x = 0.09, y = 30, width = 0.5, fontsize = 3, color = "darkgray")
  

```

# adding facet panel for frequency
```{r}
# facet_plot(tree, panel="N hits", data=subset(otu_all_table, otu_all_table$Var1 %in% otu_hits_table$Var1), geom_col, mapping = aes(x=updated_freq))

# subset(otu_all_table, Var1 %in% otu_hits_table$Var1)
frequency_barplot <-
ggplot(data = subset(otu_all_table, Var1 %in% otu_hits_table$Var1)) +
  geom_col(mapping = aes(x=updated_freq, y = Var1))+
  scale_x_reverse() +
  # theme(axis.text.y=element_blank(),
  #       axis.ticks.y=element_blank(),
  #       axis.title.y=element_blank()) +
  xlab("n Hits")
#   
# 
# tree_and_bar <- 
tree + frequency_barplot
# facet_widths(tree_and_bar, widths = c(1, 2))

```

# label leaves w/ unclassified bacteria on tree
```{r}
# unclassified_nodes <- c("Rice_field_soil_clone_Bu-23__JN615148", "uncultured_bacterium__EF645666", "uncultured_bacterium__EF645667", "uncultured_bacterium__EF645664", "uncultured_bacterium__EF645665", "uncultured_bacterium__EF645666", "Bioreactor_clone__UncS2028", "Termite_protist_endosymbiont_community_metagenome__entry118", "Polluted_aquifer_clone__UncS1623", "MSW_digester_clone__Unc01cxc", "Anaerobic_bacterium_sk.prop8__Bv2Prop2", "Anaerobic_bacterium_EtOH8__Bv2EtOH2", "Deep_sea_sediment_clone__JN798931", "Bioreactor_clone__UncS2029", "Arctic_sediment_clone_AH5_5__AtcSed67", "Pearl_river_estuary_clone_TopDSR14__UncS2036", "Polluted_aquifer_clone_LGWK16__UncS1633")

for (node in unclassified_nodes) {
  tree <- tree +  geom_cladelabel(node = MRCA(hits_only_tree, node), label = "unclassified bacteria", fontsize = 2, color = "dark grey")
}

tree
```

# tree w/ all hits
```{r}
# you got this!
# "you go girl" -Zack

# first, let's visualize the tree with all closest reference leaves and hit sequences

# load result tree in and root

# result_tree = read.tree("/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/main/workflow/out/raxmlOutputFinal/RAxML_labelledTree_noBootstrap_rooted.newick")
# rooted_result_tree <-root(result_tree, c("Moorella_thermoacetica_Y72_copy_2__MooTher5"), r = TRUE)
# # rooted_result_tree <- ggtree(rooted_result_tree, layout = "circular")
# 
# 
# # only keep leaves that are in the closest ref leaves list OR have "query" in the name
# closest_ref_leaves <- as.vector(otu_hits_table$Var1)  # closest ref leaves from otu_hits_table (made from the ref tree)
# query_leaves <- as.vector(subset(rooted_result_tree$tip.label, grepl("QUERY", rooted_result_tree$tip.label) == TRUE))
# trimmed_result_tree <- keep.tip(rooted_result_tree, c(closest_ref_leaves, query_leaves))
# trimmed_result_tree <- drop.tip(trimmed_result_tree, c(moorella_queries, "Moorella_thermoacetica_Y72_copy_2__MooTher5"))

# collapse Bilophila queries


tree <- tree_subset(trimmed_result_tree, "Bilophila_wadsworthia__BilWads6", levels_back = 4)

tree <- ggtree(tree, layout = "circular") + 
  geom_tiplab2(aes(subset= label %in% closest_ref_leaves), size = 2)
  # geom_tiplab()

tree +
  # collapse(node = MRCA(tree, bilophila_node)) %>%
  # collapse(node = MRCA(tree, dpiger_node)) +
  # collapse(node = MRCA(tree, c("QUERY___S1768", "QUERY___S2"))) +
  # collapse(node = MRCA(tree, c("QUERY___S9", "QUERY___S1286"))) +
  # collapse(node = MRCA(tree, c("QUERY___S9", "QUERY___S1286"))) +

  geom_hilight(node = MRCA(tree, bilophila_node)) +
  geom_cladelabel(node = MRCA(tree, bilophila_node), label = "Bilophila", size = 2) +
  geom_cladelabel(node = MRCA(tree, dpiger_node), label = "D. piger", size = 2)

# tree %<+%
#   otu_all_table +
#   geom_point2(aes(color = updated_freq, subset = updated_freq > 0)) +
#   scale_color_gradient(low = "blue", high = "orange") +
#   # guides(color = "none") +
#   theme(legend.position = c(0.1, 0.85))

```

# scoring system for hits with multiple closest reference sequences
```{r}
otu_hits_table <- as.data.frame(table(subset(closestRefInfo_allScoreThresholdHits_singleRefOnly$closest_ref, closestRefInfo_allScoreThresholdHits_singleRefOnly$distance < 2.5)))

# update frequencies for multiple hits
for (otu_list in otu_hits_table$Var1) {
  list_freq <- otu_hits_table$Freq[otu_hits_table$Var1 == otu_list]
  otu_list <- gsub("\\[|\\]|\\'|\\ ", "", otu_list)
  otu_list <- strsplit(otu_list, ",")[[1]]
  if (length(otu_list) > 1) {
    for (otu in otu_list){
      freq <- otu_hits_table$Freq[gsub("\\[|\\]|\\'|\\ ", "",otu_hits_table$Var1) == otu]
      added_freq <- list_freq / length(otu_list)
      updated_freq <- freq + added_freq
      otu_hits_table$updated_freq[gsub("\\[|\\]|\\'|\\ ", "",otu_hits_table$Var1) == otu] <- updated_freq
    }
  }
}

# add Freq to updated frequencies for otus w/o multiple hits
for (otu_list in otu_hits_table$Var1[is.na(otu_hits_table$updated_freq)]) {
  list_freq <- otu_hits_table$Freq[otu_hits_table$Var1 == otu_list]
  otu_list <- gsub("\\[|\\]|\\'|\\ ", "", otu_list)
  otu_list <- strsplit(otu_list, ",")[[1]]
  if (length(otu_list) == 1) {
    otu_hits_table$updated_freq[gsub("\\[|\\]|\\'|\\ ", "",otu_hits_table$Var1) == otu_list] <- list_freq
  }
}
# make new df and format otu name
otu_hits_table <- subset(otu_hits_table, !is.na(otu_hits_table$updated_freq))
otu_hits_table$Var1 <- gsub("\\[|\\]|\\'|\\ ", "", otu_hits_table$Var1)

```

# remove Moorella thermoacetica hits from df and update otu_hits and tree
```{r}
# closest_refInfo_noMThermCopy2 <- subset(closestRefInfo_allScoreThresholdHits_singleRefOnly, grepl("copy_2", closestRefInfo_allScoreThresholdHits_singleRefOnly$closest_ref) == FALSE)
# moorella_queries <- queryDistanceInfo$query_names[grepl("copy_2", queryDistanceInfo$closest_ref) == TRUE]
# 
# ggplot(subset(closestRefInfo_allScoreThresholdHits_singleRefOnly, grepl("Archaeoglobus", closestRefInfo_allScoreThresholdHits_singleRefOnly$closest_ref) == TRUE)) +
#   geom_histogram(aes(x = distance))

tree <- phyloseq(trimmed_result_tree)
# moorella_clade <- tree_subset(tree, node = MRCA(tree, c("Moorella_thermoacetica_Y72_copy_2__MooTher5", "QUERY___S1460")), levels_back = 0)
# clade_level2 <- tree_subset(tree, node = MRCA(tree, c("Moorella_thermoacetica_Y72_copy_2__MooTher5", "QUERY___S1460")), levels_back = 2)
# 
# ggtree(clade_level2, aes(color = group)) + geom_tiplab()

# confirm that moorella_clade identifies all queries in the copy 2 clade
# now let's remove it from the tree

tree <- drop.tip(tree, moorella_clade$tip.label)

ggtree(tree, layout = "circular") +
  geom_label_repel(aes(label= ifelse(label %in% closest_ref_leaves, label, NA)), size = 1, max.overlaps = Inf, label.padding = 0.1)
```

# visualizing result tree with phyloseq and tip_glom
```{r}
# prep the tree: read in file, root, and trim to include only closest ref leaves and query leaves
# result_tree = read.tree("/Users/rebeccachristensen/Desktop/Cremer_Lab_2022/main/workflow/out/raxmlOutputFinal/RAxML_labelledTree_noBootstrap_rooted.newick")
# rooted_result_tree <-root(result_tree, c("Moorella_thermoacetica_Y72_copy_2__MooTher5"), r = TRUE)
# closest_ref_leaves <- as.vector(otu_hits_table$Var1)  # closest ref leaves from otu_hits_table (made from the ref tree)
# query_leaves <- as.vector(subset(result_tree$tip.label, grepl("QUERY", result_tree$tip.label) == TRUE))
# trimmed_result_tree <- keep.tip(result_tree, c(closest_ref_leaves, query_leaves))
# 
# # remove moorella copy 2 from the tree
# moorella_clade <- tree_subset(trimmed_result_tree, node = MRCA(trimmed_result_tree, c("Moorella_thermoacetica_Y72_copy_2__MooTher5", "QUERY___S1460")), levels_back = 0)
# trimmed_result_tree <- drop.tip(trimmed_result_tree, moorella_clade$tip.label)

# 
tree <- phyloseq(trimmed_result_tree)
# 

# subsetting by Bilophila
# tree <- tree_subset(trimmed_result_tree, "Bilophila_wadsworthia__BilWads6", levels_back = 4)
# plot_tree(subset, label.tips = "taxa_names")
# plot_tree(tip_glom(subset, h = 0.8), label.tips = "taxa_names")

# labeling reference sequences
# plot_tree(subset)

tree <- 
  ggtree(tree, layout = "circular") +
  # geom_label_repel(aes(label= ifelse(label %in% closest_ref_leaves, label, NA)), size = 1, max.overlaps = Inf, label.padding = 0.1, nudge_x = 2)
  geom_tiplab2(aes(subset=(label %in% closest_ref_leaves)), align = TRUE, size = 2)

tree %<+%
  data +
  aes(color = group) +
  scale_color_manual(values = c("red", "blue", "orange")) +
  guides(color = "none")

```


# categorize leaves as query or ref (for aesthetic mapping)
```{r}
tree <- phyloseq(trimmed_result_tree)
data <- data_frame("tip"= tree$tip.label)
data$ref_or_query <- ifelse(grepl("QUERY", data$tip), "query", "ref")

desulfovibronaceae_clade <- tree_subset(tree, node = MRCA(tree, desulfovibrionaceae_node), levels_back = 0)
firmicutes_clade <- tree_subset(tree, node = MRCA(tree, firmicutes_node), levels_back = 0)
# archaeoglobus_clade <- tree_subset(tree, node = MRCA(tree, archaeoglobus_node), levels_back = 2)

data$group <- ifelse(data$tip %in% desulfovibronaceae_clade$tip.label, "Desulfovibrionaceae", 
                     ifelse(data$tip %in% firmicutes_clade$tip.label, "Firmicutes group", "other"))
```
